---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)
```

# mongooser

<!-- badges: start -->
<!-- badges: end -->

[WIP / DO NOT USE IN PRODUCTION]

The goal of mongooser is to port [mongoosejs]() to R, for a more, predictable use of MongoDB in a production context.

## Installation

You can install the development version of mongooser like so:

``` r
remotes::install_github("thinkr-open/mongooser")
```

### What's wrong with `{mongolite}`?

Nothing, `{mongooser}` is actually built on top of it. 

### Ok, so why?

An issue with MongoDB is that it doesn't enforce any kind of schema, and R isn't very well equiped to handle unstructured data. 

Also, given that it's unstructured and does the type conversion for you, you can get unexpected results when using MongoDB with R. 

And on top of that, human tend to be a little bit lazy, and if the db allows them to insert random unstructured elements in the db, well, they'll do it, and then some data engineer will cry and scream to scrap and restructure the data to be usable. 

Here is a simple example:

```{r}
con <- mongolite::mongo()
con$insert(data.frame(x = 1))
con$insert(data.frame(x = "this"))
res <- con$find()
class(res$x)
res <- con$find('{"x":1}')
class(res$x)
res <- con$find('{"x":"this"}')
class(res$x)
```

As MongoDB & `{mongolite}` doesn't inforce any kind of type on read/write, it can be hard to rely on the output. 

It's even more complex when you start having nested elements. 

```{r}
con$drop()
con <- mongolite::mongo()
con$insert(
  list(
    y = data.frame(x = 1)
  )
)
# You'd expect to have a list with a data.frame
con$find()
con$insert(
  list(
    z = data.frame(x = 1)
  )
)
# You'd expect to have 2 lists with a data.frame
con$find()

# You can also do
it <- con$iterate()
while (!is.null(x <- it$one())) {
  print(x)
}
# But that's still not exactly what you'd want
```

`{mongooser}` tries to solve that issue by porting [mongoose.js]() to R. 

## Here is an example 

```
docker run -d -p 2811:27017 mongo:3.4
```

```{r error = TRUE}
library(mongooser)
# New instance of Mongoose
mongoose <- Mongoose$new()

# # Connect to the DB
mongoose$connect(
  url = "mongodb://localhost:2811"
)

# This create an object that will
# query / write to the `name`
# collection (here Cat)
Cat <- mongoose$model(
  name = "Cat",
  schema = list(
    name = list(
      type = "character"
    )
  )
)
# You can drop everything from the collection if you want
Cat$drop()

# Cat$new() will create a new document object,
# You __need__ to pass to the object data that
# match the schema defined in the model

# This should fail => bad type
kitty <- Cat$new(
  list(
    name = 1234
  )
)
# This should fail => undefined schema entry
kitty <- Cat$new(
  list(
    name = 1234,
    weight = 12
  )
)

# This will work
fluffy <- Cat$new(
  list(
    name = "Fluffy"
  )
)

# Expected output => An empty list
Cat$find()

# Saving the fluffy doc
fluffy$save()

# Expected output =>
# A list of length one, and the first element is
# $name
# [1] "Fluffy"
Cat$find()

# Let's now create a new document
minette <- Cat$new(
  list(
    name = "minette"
  )
)
# Save it
minette$save()

# Expected output =>
# A list of length two, with both elements being
# $name
# [1] "" <- character string
Cat$find()

# Let's now compare to what you'd get with
# base mongolite
con <- mongolite::mongo(collection = "Cat")
con$drop()
con$insert(
  list(
    name = "Fluffy"
  )
)
con$find()
con$insert(
  list(
    name = "minette"
  )
)
con$find()
```

Here for a more complex example with a dataframe inside
```{r}
Cat <- mongoose$model(
  name = "Cat",
  schema = list(
    name = list(
      type = "character"
    ),
    food = list(
      type = "data.frame"
    )
  )
)

Cat$drop()

fluffy <- Cat$new(
  list(
    name = "fluffy",
    food = data.frame(
      date = "2022-02-12",
      type = "fish"
    )
  )
)
fluffy$save()

# Expected output => list of length one,
# with one character string in "name" and
# one data frame in "food"
Cat$find()

minette <- Cat$new(
  list(
    name = "minette",
    food = data.frame(
      date = "2022-02-12",
      type = "cake"
    )
  )
)
minette$save()

# Expected output => list of length two,
# each element has
# one character string in "name" and
# one data frame in "food"
Cat$find()

# Let's now compare to mongolite default behavior
con <- mongolite::mongo(collection = "Cat")
con$drop()
con$insert(
  list(
    name = "fluffy",
    food = data.frame(
      date = "2022-02-12",
      type = "fish"
    )
  )
)
con$find()

con$insert(
  list(
    name = "minette",
    food = data.frame(
      date = "2022-02-12",
      type = "cake"
    )
  )
)
con$find()
```